---
title: "RSCR paper code: Introductory figure"
author: "Joshua Cape (jrcape@wisc.edu)"
output: html_notebook
---

## Preamble
```{r message=FALSE}
rm(list = ls())
source("RSCR_preamble.R")

mix_func <- function(mu,sd,eps){
  alloc <- runif(1) < 1-eps
  return(alloc*rnorm(1,mean=mu,sd=sd) + (!alloc)*rnorm(1,mu,100*sd))}
  
library(knitr)
```


## Blockwise contaminated normal data
NB: naively generates IID entries, then symmetrizes.
```{r}
set.seed(0)

temp.eps <- 0.01

temp.n <- 1000
temp.props <- c(1/2,1/2)
temp.B <- rbind(c(2,1),c(1,2))
temp.sd <- 3
# (automated below)
temp.memb <- sapply(1:length(temp.props),
                    function(xxx) t(replicate(temp.props[xxx]*temp.n, diag(length(temp.props))[xxx,])))
temp.K <- length(temp.props)
temp.pop.mtx <- temp.memb %*% temp.B %*% t(temp.memb)
temp.data.raw <- matrix(sapply(temp.pop.mtx, function(xxx) mix_func(xxx,temp.sd,temp.eps)), nrow=temp.n)
  
data.mtx <- temp.data.raw
data.mtx[lower.tri(data.mtx,diag=TRUE)] <- 0
data.mtx <- data.mtx + t(data.mtx)

dim.mtx <- temp.n
mu.1 <- temp.B[1,1]
mu.2 <- temp.B[1,2]
mu.3 <- temp.B[2,2]
eps <- temp.eps
```


## Eigenvector computation
(i) noiseless ground truth, (ii) raw data, (iii) ptr matrix
```{r}
temp.range <- c(-2,2)

    pop.vec1 <- c(rep(1,dim.mtx/2),rep(1,dim.mtx/2))
    pop.vec2 <- c(rep(1,dim.mtx/2),rep(-1,dim.mtx/2))

my_signed_median <- function(input){return(input*sign(median(input)))} # sign flip via median
my_signed_vec2 <- function(input){return(input*sign(input*pop.vec2))}
  
  embed.raw <- sqrt(dim.mtx)*eigs(data.mtx,4,"LR")$vectors
  
  embed.ptr <- sqrt(dim.mtx)*eigs(ptr_func(data.mtx),2,"LR")$vectors
  embed.ptr[,1] <- my_signed_median(embed.ptr[,1])
  embed.ptr[,2] <- my_signed_vec2(embed.ptr[,2])
  
my.ylab <- "Value"
my.xlab1 <- "Entry (node index)"
my.xlab2 <- "Entry (node index)"
my.pch <- 20
my.col.pts <- "gray"
my.col.l <- "blue"
my.lwd <- 4
```


## Eigenvalue computation - method selects largest real part
```{r}
set.length = 10
  
  evals.pop <- eigs(temp.pop.mtx, set.length, "LR")$values
  evals.raw <- eigs(data.mtx, set.length, "LR")$values
  evals.ptr <- eigs(ptr_func(data.mtx), set.length, "LR")$values

mys.xlab <- "Eigenvalue index"
mys.ylab <- "Value"
mys.pch <- 20
mys.lwd <- 3
mys.col.pts <- "gray"
mys.col.l <- "black"
mys.cex <- 2

mys.cex.axis <- 1.5
mys.cex.lab <- 1.5
mys.cex.main <- 1.5
```


## Save and print - scree plots
```{r}
pdf(file="fig_intro_scree.pdf", width=8, height=4)
par(mfrow = c(1, 2))
plot(evals.raw/dim.mtx, ylim=c(-0.0001,max(evals.raw/dim.mtx)), main="Scree plot of data matrix",
xlab=mys.xlab, ylab=mys.ylab, pch=mys.pch, col=mys.col.pts, cex=mys.cex, las=1, cex.axis=mys.cex.axis, cex.lab=mys.cex.lab, cex.main=mys.cex.main)
abline(h=0, col=mys.col.l, lwd=mys.lwd)
    
plot(evals.ptr/dim.mtx, ylim=c(-0.0001,max(evals.ptr/dim.mtx)), main="Scree plot of PTR matrix",
xlab=mys.xlab, ylab=mys.ylab, pch=mys.pch, col=mys.col.pts, cex=mys.cex, las=1, cex.axis=mys.cex.axis, cex.lab=mys.cex.lab, cex.main=mys.cex.main)
abline(h=0, col=mys.col.l, lwd=mys.lwd)
dev.off()
#plot_crop("fig_intro_scree.pdf")
```


### Save and print - top eigenvector, raw and ptr side-by-side
```{r message=FALSE}
pdf(file="fig_intro_evec1.pdf", width=8, height=4)
par(mfrow = c(1, 2))
plot(my_signed_median(embed.raw[,1]), main="1st eigenvector of data matrix", ylim=temp.range, ylab=my.ylab, xlab=my.xlab1, pch=my.pch, col=my.col.pts, las=1, cex.axis=mys.cex.axis, cex.lab=mys.cex.lab, cex.main=mys.cex.main)
lines(1:dim.mtx,rep(1,dim.mtx),type="l",col=my.col.l,lwd=my.lwd)    

plot(embed.ptr[,1], main="1st eigenvector of PTR matrix", ylim=temp.range, ylab=my.ylab, xlab=my.xlab1, pch=my.pch, col=my.col.pts, las=1, cex.axis=mys.cex.axis, cex.lab=mys.cex.lab, cex.main=mys.cex.main)
lines(1:dim.mtx,rep(1,dim.mtx),type="l",col=my.col.l,lwd=my.lwd)   
dev.off()
```


### Save and print - second eigenvector, raw and ptr side-by-side
```{r}
pdf(file="fig_intro_evec2.pdf", width=8, height=4)
par(mfrow = c(1, 2))
plot(my_signed_median(embed.raw[,2]), main="2nd eigenvector of data matrix", ylim=temp.range, ylab=my.ylab, xlab=my.xlab2, pch=my.pch, col=my.col.pts, las=1, cex.axis=mys.cex.axis, cex.lab=mys.cex.lab, cex.main=mys.cex.main)
lines(1:(dim.mtx/2),rep(1,dim.mtx/2),type="l",col=my.col.l,lwd=my.lwd)
lines((dim.mtx/2+1):dim.mtx,rep(-1,dim.mtx/2),type="l",col=my.col.l,lwd=my.lwd)

plot(embed.ptr[,2], main="2nd eigenvector of PTR matrix", ylim=temp.range, ylab=my.ylab, xlab=my.xlab2, pch=my.pch, col=my.col.pts, las=1, cex.axis=mys.cex.axis, cex.lab=mys.cex.lab, cex.main=mys.cex.main)
lines(1:(dim.mtx/2),rep(1,dim.mtx/2),type="l",col=my.col.l,lwd=my.lwd)
lines((dim.mtx/2+1):dim.mtx,rep(-1,dim.mtx/2),type="l",col=my.col.l,lwd=my.lwd)
dev.off()
```
