---
title: "RSCR paper code: Connectome data"
author: "Joshua Cape (jrcape@wisc.edu)"
output: html_notebook
---

## Preamble, load data, and remove scanner error graph
```{r message=FALSE}
rm(list = ls())
source("RSCR_preamble.R")
# cf. https://github.com/youngser/TwoTruth

import.dataset.name <- "DS01876"

print(
  load(
    url(
      paste("http://www.cis.jhu.edu/~parky/TT/Data/TT-DS",import.dataset.name,"-glist114-raw.rda", sep=""))))

#Below: remove scanner error graph
xxx <- unlist(lapply(glist, vcount)); which(xxx == min(xxx)); rm(xxx)
glist <- glist[-107]; # find and remove technical error graph
```


```{r}
# Sanity check: each graph is a single connected component
table(unlist(lapply(glist, is_connected))) # TRUE, 113
table(unlist(lapply(glist, count_components))) # 1, 113
```


## Description - graph vertices
```{r}
# Per-graph vertex counts
df.vcounts <- melt(data.frame(vcounts = unlist(lapply(glist, vcount))))

# Per-graph median vertex strength
df.med.str <- melt(data.frame(med.str = sapply(1:length(glist), function(xxx) median(strength(glist[[xxx]])))))

# Per-graph maximum vertex strength
df.max.str <- melt(data.frame(max.str = sapply(1:length(glist), function(xxx) max(strength(glist[[xxx]])))))

my_plotter <- function(my.df, my.x.lab){
  ggplot(my.df, aes(x=value, colour=variable)) +
    geom_histogram(bins=10) +
    ylim(c(0,30)) + xlab(my.x.lab) + ylab("Graph frequency") +
    theme(legend.position = "none",
          legend.text = element_text(size=16),
          axis.text.x = element_text(size=16, vjust = 0.75, hjust = 0.75),
          axis.text.y = element_text(size=16),
          axis.title = element_text(size=16, face="bold"),
          plot.margin = grid::unit(c(0,0,0,0), "mm"),
          aspect.ratio = 1) #+ scale_x_continuous(labels = function(xxx) format(xxx, scientific = TRUE))
}

my_plotter(df.vcounts, "Vertex count")
```


## Save output
```{r}
ggsave("fig_ptr_brains_vcounts.pdf",
       my_plotter(df.vcounts, "Vertex count"),
       width = 3.5,
       height = 3.5)
```


## Table of summary statistics
```{r}
# Per-graph summary statistics
func_summary <-
  function(xxx){return(c(median(xxx), mean(xxx), max(xxx)))}

# Edge weight summary statistics -- median over all graphs
edge.summary.mtx <-
  matrix(do.call(rbind, lapply(glist,function(xxx) func_summary(E(xxx)$weight))), nrow = length(glist))

colMedians(edge.summary.mtx)

# Vertex strength summary statistics -- median over all graphs
strength.summary.mtx <-
  matrix(do.call(rbind, lapply(glist, function(xxx) func_summary(strength(xxx))
)), nrow = length(glist))

colMedians(strength.summary.mtx)
```


## Description - box plots of vertex strengths for several graphs
```{r}
num.graphs <- 20 # >> if all desired, then use length(glist)

max.vcount <- max(unlist(lapply(glist, vcount)))

graph.strengths <- lapply(1:num.graphs,
                          function(xxx) strength(glist[[xxx]]))
graph.strengths <- lapply(1:num.graphs,
                          function(yyy) append(graph.strengths[[yyy]],
                                               rep(NA, max.vcount - length(graph.strengths[[yyy]]) + 1)))
graph.strengths <-
  do.call("cbind", lapply(1:num.graphs, function(yyy) graph.strengths[[yyy]]))
               
my.df <- melt(graph.strengths)

my.boxplot.strengths <- 
ggplot(my.df, aes(x=factor(Var2),y=value), asp=1) +
  geom_boxplot(outlier.size = 0.5) +
  xlab("Index of graph") +
  ylab("Vertex strength distribution") +
  scale_x_discrete(limits=paste(seq(1,num.graphs))) +
  theme(legend.position = "none",
        legend.text = element_text(size=16),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=16),
        axis.title = element_text(size=16, face="bold"),
        plot.margin = grid::unit(c(0,0,0,0), "mm")) #+ coord_fixed(ratio=num.graphs/max(na.omit(my.df$value))/2)

my.boxplot.strengths
```


## Save output
```{r}
ggsave("fig_ptr_brains_vstrengths.pdf",
       my.boxplot.strengths,
       width = 7,
       height = 3.5)
```


## (approx 8 minutes) Compute Zhu--Ghodsi embedding dimension values
We apply the Zhu--Ghodsi (ZG) method for dimension selection. We apply the procedure to the top 50 singular values of each data set, dropping the leading (largest) value in the interest of mitigating possible shadowing effects. At the same time, in an effort to compensate for possible over-selection, we leave the ZG value unaltered.
```{r message=FALSE}
start_time <- Sys.time() # <--------------------

num.scree <- 50 # max scree index

set.seed(1234)

my.ZG.vals <-
  lapply(glist,
         function(xxx)
           c(dim_select(svdr(xxx[TRUE,TRUE], k=num.scree)$d[-1]),
             dim_select(svdr(ptr_func(xxx[TRUE,TRUE]), k=num.scree)$d[-1])))

my.ZG.vals <- do.call(rbind,my.ZG.vals)

table(my.ZG.vals[,1] - my.ZG.vals[,2])

end_time <- Sys.time() # <----------------------
end_time - start_time
system("say Done")
```


## Compare ZG values between raw data embedding and ptr embedding
```{r}
temp.df <- data.frame(my.ZG.vals)

plot.raw.vs.ptr.embed.dim <-
  ggplot(temp.df, aes(X1, X2)) +
  geom_count(aes(size = ..n..)) +
  geom_abline(slope=1, linetype="dashed") +
  scale_size_continuous(name = "counts") +
  xlim(min(my.ZG.vals), max(my.ZG.vals)) +
  ylim(min(my.ZG.vals), max(my.ZG.vals)) +
  xlab("Raw embedding dimension") +
  ylab("PTR embedding dimension") +
  coord_fixed() +
     theme(legend.justification = c(1, 1),
          legend.position = "none",
          legend.key.size = unit(1, 'cm'),
          legend.text = element_text(size=16),
          legend.title = element_blank(),
          axis.text.x = element_text(size=16),
          axis.text.y = element_text(size=16),
          axis.title=element_text(size=16,face="bold"))

plot.raw.vs.ptr.embed.dim

rm(temp.df)
```


### Generate pdf output image
```{r}
ggsave("fig_ptr_brains_dhats.pdf",
       plot.raw.vs.ptr.embed.dim,
       width = 3.5,
       height = 3.5)
```


## Save truncated eigendecompositions of A and PTR(A)
```{r message=FALSE}
start_time <- Sys.time() # <-----------------------

  # >> uses input "my.ZG.vals"

  num.true.clust <- 4

  raw.evec.list <- list()
  ptr.evec.list <- list()
  
  set.seed(1234)
  
  for(g.index in 1:length(glist)){
      my.graph <- glist[[g.index]]
      my.adj <- my.graph[TRUE,TRUE]
      
      raw.evec.list[[g.index]] <- svdr(my.adj, k = my.ZG.vals[g.index,1])$u
      ptr.evec.list[[g.index]] <- svdr(ptr_func(my.adj), k = my.ZG.vals[g.index,2])$u
     
  }
  
end_time <- Sys.time() # <-----------------------

end_time - start_time # <-----------------------

system("say Done")

```


## Plot average ARI across replicates of k-means clustering
```{r}

set.seed(1234)
n.reps <- 100

raw.evec.kmean.ARI.data.func <- function(yyy)
  lapply(1:length(glist),
         function(xxx)
           adjustedRandIndex(kmeans(raw.evec.list[[xxx]], num.true.clust)$cluster,
                             V(glist[[xxx]])$Y))
ptr.evec.kmean.ARI.data.func <- function(yyy)
  lapply(1:length(glist),
         function(xxx)
           adjustedRandIndex(kmeans(ptr.evec.list[[xxx]], num.true.clust)$cluster,
                             V(glist[[xxx]])$Y))

raw.evec.kmean.ARI.data <- lapply(1:n.reps, raw.evec.kmean.ARI.data.func)
ptr.evec.kmean.ARI.data <- lapply(1:n.reps, ptr.evec.kmean.ARI.data.func)

raw.evec.kmean.ARI.traj <-
  colMeans(matrix(unlist(do.call(rbind, raw.evec.kmean.ARI.data)), nrow = n.reps))
ptr.evec.kmean.ARI.traj <-
  colMeans(matrix(unlist(do.call(rbind, ptr.evec.kmean.ARI.data)), nrow = n.reps))


plot(ptr.evec.kmean.ARI.traj, ylim = c(0,0.2),
     type="l",
     col="blue",
     xlab = "Index of graph",
     ylab = "Average ARI")
lines(raw.evec.kmean.ARI.traj, ylim = c(0,0.2),
      type="l",
      col="red")



my.data.list <- list()

for(iii in 1:length(glist)){
      my.data.list[[iii]] <-
        c(iii,
          raw.evec.kmean.ARI.traj[iii],
          ptr.evec.kmean.ARI.traj[iii])
      names(my.data.list[[iii]]) <-
        c('g.index',
          'raw.km.ARI',
          'ptr.km.ARI')
}

my.ARI.data <- as.data.frame(do.call("rbind", lapply(my.data.list, function(xxx) xxx)))
my.ARI.data <- melt(my.ARI.data, id.vars = 'g.index', variable.name = 'series')

plot.raw.vs.ptr.brains.ARI.km <-
  ggplot(my.ARI.data, aes(g.index, value)) +
  geom_line(aes(colour = series), size=1) +
  scale_color_manual(labels=c("Raw, k-means","PTR, k-means"), values=ggplotColours(2)) +
  scale_x_discrete(limits=c(1,seq(10,length(glist),by=10))) + ylim(0,0.20) +
  xlab("Index of graph") + ylab("Adjusted Rand index") +
   theme(legend.justification = c(1, 1),
        legend.position = c(1,1),
        legend.direction = "horizontal",
        legend.key.size = unit(1, 'cm'),
        legend.text = element_text(size=16),
        legend.title = element_blank(),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=16),
        axis.title=element_text(size=16,face="bold"))

plot.raw.vs.ptr.brains.ARI.km

```
## Save plot of average ARI
```{r}
ggsave("fig_ptr_brains_ARI_kmeans.pdf",
       plot.raw.vs.ptr.brains.ARI.km,
       width = 7,
       height = 3.5)
```

